A Framework for Discrete Change

### 2008-11-06 papers/individual/individual.tex
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/depmix@236 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

The Weather Prediction Task (WPT, Knowlton, Squire \& Gluck, 1994) 
is a probabilistic categorization task, in which participants learn to 
predict the state of the weather (sunny, or rainy) on the basis 
of four ``tarot'' cards (cards with abstract geometrical patterns). 
Each cue pattern is associated with a particular probability distribution 
over the states of the weather. In order to perform in the task, 
participants must predict the weather in accordance with these
conditional probabilities.

@@

As we fit the data to a single subject, 
we must place some constraints. Specifically, we constrain the state 
transitions to be in a ``left-right'' format (states can only proceed 
to the immediately adjacent state and never back, and must start in 
the initial state). We fitted a single, two and three state model to 
the data. This showed that a two state model was better than a single
and three state model.

@@

Investigation of the parameter estimates (see Table~\ref{tab:WPT}) indicated that the first state might be
a single cue strategy (the regression coefficient for the first cue was of much 
larger magnitude than that of the other cues). The second state was a multi-cue
strategy (all cues had regression coefficients of reasonable magnitude). 
To reduce the degrees of freedom, and improve parameter estimates, we implemented 
constraints to force state 1 into a single cue strategy (fixing the coefficients
of the remaining three cues to 0) and state 2 in a multi-cue strategy (forcing
the intercept to 0). These restrictions resulted in a better AIC value of AIC=185.24 
(df=7). Interestingly, the single cue strategy was somewhat different than 
described by Gluck et al. Parameter estimates indicated relatively more consistent 
predictions of ``rain'' in the absence of cue 1 ($Pr(\text{sun}) = 0.22$) and more
inconsistent predictions of ``sun'' in the presence of cue 1 ($Pr(\text{sun}) = 0.60$). 
The cue weights of the multi-cue strategy were in the direction of the optimal weights. 
The Viterbi state sequence indicated that the participant used the single
cue strategy for the first 60 trials, and then switched to the multi-cue strategy.

>>>>

The Weather Prediction Task (WPT, Knowlton, Squire \& Gluck, 1994)
is a probabilistic categorization task in which participants learn to
predict (or categorize) the state of the weather (sun or rain) on the basis
of four ``tarot'' cards (cards with abstract geometrical patterns). On a
given trial, one, two, or three cues are present. There are a total of
14 possible cue patterns, and each one is associated with a particular
probability distribution over the states of the weather. In order to
perform in the task, participants must predict the weather in accordance
with these conditional probabilities.

@@

As we fitted a DMM to the data of a single subject, it was necessary
to place some constraints on the model. Specifically, we constrain the state
transitions to be in a ``left-right'' format (states can only proceed
to the immediately adjacent state and never back, and must start in
the initial state). We fitted a single, two and three state model to
the data. This showed that a two state model was better than a single
or three state model.

@@

Investigation of the parameter estimates (see Table~\ref{tab:WPT}) showed that
the regression coefficient for the first cue was of much
larger magnitude than that of the other cues. As such, the first
state seems representative of single cue strategy. Alternatively,
as all regression coefficients are positive, it could indicate a ``counting''
heuristic, where the propensity of ``sun'' responses increases when
more cues are present (regardless of which cues they are). However, in that
case, we would expect the regression coefficients to be of roughly identical
magnitude. The second state seemed to represent a multi-cue strategy, as all
cues had regression coefficients of reasonable magnitude, with differing
directions in line with the objective cue validities.

To reduce the degrees of freedom, and improve parameter estimates, we
implemented constraints to force state 1 into a single cue strategy (fixing the
coefficients of the remaining three cues to 0) and state 2 in a multi-cue
strategy (forcing the intercept to 0). These restrictions resulted in a better
AIC value of AIC=185.24 (df=7). Interestingly, the single cue strategy was
somewhat different than described by Gluck et al. Parameter estimates indicated
relatively more consistent predictions of ``rain'' in the absence of cue 1
($Pr(\text{sun}) = 0.22$) and more inconsistent predictions of ``sun'' in the
presence of cue 1 ($Pr(\text{sun}) = 0.60$). The cue weights of the multi-cue
strategy were in the direction of the optimal weights. The Viterbi state
sequence indicated that the participant used the single cue strategy for the
first 60 trials, and then switched to the multi-cue strategy.
